---
title: "AMS 580 project Kaggle: Spaceship Titanic, using Support Vector Machine"
author: "580 Group Project (YJ,JY,BL,YG)"
date: "2022/5/10"
output: word_document
---

```{r}
if (!requireNamespace("tidyverse")) install.packages('tidyverse')
if (!requireNamespace("caret")) install.packages('caret')

library(tidyverse)
library(caret)
```

Read Training data set and divide the cleaned data into 90% training and 10% testing.
```{r}
train.data <- read.csv('~/Desktop/Spring 2022/AMS 580/580 group proj/spaceship-titanic/spcshipTtn-train.csv')
train.data <- subset(train.data, select = -c(PassengerId, Cabin, Name))
train.data <- na.omit(train.data)

train.data$CryoSleep = ifelse(train.data$CryoSleep == "True", 1, 0)
train.data$VIP = ifelse(train.data$VIP == "True", 1, 0)
train.data$Transported = ifelse(train.data$Transported == "True", 1, 0)

train.data$CryoSleep<- as.factor(train.data$CryoSleep)
train.data$VIP<- as.factor(train.data$VIP)
train.data$Transported <- as.factor(train.data$Transported)

set.seed(123)
training.samples <- train.data$Transported %>%
  createDataPartition(p = 0.90, list = FALSE)
Ttrain.data  <- train.data[training.samples, ]
Ttest.data <- train.data[-training.samples, ]

str(Ttrain.data) # 6859 obs
str(Ttest.data) # 761 obs
```

Build the best classifier to predict whether a check is forged or not using the training data and the linear SVM. 
Contains confusion matrix and the sensitivity. 
Finally, report the overall accuracy.
```{r}
# Fit the model on the training set
set.seed(123)
model <- train(
  Transported ~., data = Ttrain.data, method = "svmLinear",
  trControl = trainControl("cv", number = 10),
  # This will scale all the predictors in the model
  )

# Make predictions on the test data
predicted.classes <- model %>% predict(Ttest.data)
head(predicted.classes)

# Confusion matrix
confusionMatrix(factor(predicted.classes), factor(Ttest.data$Transported), positive = '1')

# Accuracy 0.8133
```

Build the best classifier to predict whether a check is forged or not using the training data and the linear SVM. 
This time find the optimal cost parameter C by using tuneGrid = expand.grid(C = seq(0, 2, length = 20)).
Contains the plot of model accuracy vs different values of cost, confusion matrix and the sensitivity. 
Finally, report the overall accuracy.
```{r}
# Fit the model on the training set
set.seed(123)
model <- train(
  Transported ~., data = Ttrain.data, method = "svmLinear",
  trControl = trainControl("cv", number = 10),
  tuneGrid = expand.grid(C = seq(0.1, 2, length = 19))
  )

# Plot model accuracy vs different values of Cost
plot(model)

# Print the best tuning parameter C that
# maximizes model accuracy
model$bestTune

# Make predictions on the test data
predicted.classes <- model %>% predict(Ttest.data)

# Confusion matrix
confusionMatrix(factor(predicted.classes), factor(Ttest.data$Transported), positive = '1')

# Accuracy 0.8120
```

Build the best classifier to predict whether a check is forged or not using the training data and the SVM with radial basis kernel. 
This time find the optimal tuning parameters C and sigma by using the command line: tuneLength = 10.
Contains confusion matrix and the sensitivity. 
Finally, report the overall accuracy.
```{r}
# Fit the model on the training set
set.seed(123)
model <- train(
  Transported ~., data = Ttrain.data, method = "svmRadial",
  trControl = trainControl("cv", number = 10),
  tuneLength = 10
  )
# Print the best tuning parameter sigma and C that
# maximizes model accuracy
model$bestTune

# Make predictions on the test data
predicted.classes <- model %>% predict(Ttest.data)

# Confusion matrix
confusionMatrix(factor(predicted.classes), factor(Ttest.data$Transported), positive = '1')

# Accuracy 0.8054
```

Build the best classifier to predict whether a check is forged or not using the training data and the SVM with polynomial basis kernel.
This time find the optimal tuning parameters C, degree and scale by using the command line: tuneLength = 3.
Contains confusion matrix and the sensitivity. 
Finally, report the overall accuracy.
```{r}
# Fit the model on the training set
set.seed(123)
model <- train(
  Transported ~., data = Ttrain.data, method = "svmPoly",
  trControl = trainControl("cv", number = 10),
  tuneLength = 3
  )
# Print the best tuning parameter sigma and C that
# maximizes model accuracy
model$bestTune

# Make predictions on the test data
predicted.classes <- model %>% predict(Ttest.data)

# Confusion matrix
confusionMatrix(factor(predicted.classes), factor(Ttest.data$Transported), positive = '1')

# Accuracy 0.8119
```


Overall accuracy rate:

Linear: About 0.8133 (The Best) & 0.8120 (using tuneGrid)

Radial basis kernel: 0.8054

Polynomial kernel: 0.8119

SVM Linear gets the highest accuracy.
